Stochastic Variance Reduced Policy Gradient

### 2018-01-08 svrpg.tex
written convex SVRG part; some stuff will probably be moved to the introduction; may add pseudocode for SVRG

\subsection{Stochastic Variance-Reduced Gradient}
Finite sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\vtheta)$:
\begin{align*}
\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta).
\end{align*}
This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$. In this case, we adopt the following, more meaningful notation:
\begin{align*}
\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_{\vtheta}(x_i).
\end{align*}
A common requirement is that $g$ must be smooth and convex in $\vtheta$. Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an IFO complexity \cite{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$. This can be too expensive for large values of $N$. Stochastic gradient ascent \cite{robbins1951stochastic} \cite{bottou2004large} overcomes this problem, at the cost introducing some variance. The IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation: it does not depend on $N$, but it has a worse dependency on $\epsilon$.

\section{Algorithm}

>>>>

\subsection{Stochastic Variance-Reduced Gradient}
Finite sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\vtheta)$:
\begin{align*}
\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta).
\end{align*}
This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$. In this case, we adopt the following, more meaningful notation:
\begin{align*}
\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_{\vtheta}(x_i).
\end{align*}
A common requirement is that $g$ must be smooth and convex in $\vtheta$. Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an IFO complexity \cite{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$. This can be too expensive for large values of $N$. Stochastic gradient ascent \cite{robbins1951stochastic} \cite{bottou2004large} overcomes this problem by sampling a single $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling. The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$. 
Starting from SAG \cite{roux2012stochastic}, a series of variations to stochastic gradient ascent have been proposed to achieve a better trade-off between convergence speed and cost per iteration. This family of algorithms includes SAG, SVRG \cite{johnson2013accelerating}, SAGA \cite{defazio2014saga} and Finito \cite{defazio2014finito}, the common intuition being the reuse of previous gradient computations to reduce the variance of the current one. In particular, Stochastic Variance Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when neural networks with several parameters are employed.

The idea of SVRG is to alternate full and stochastic gradient updates. Each $m$ iterations, a snapshot $\tilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\tilde{\vtheta})$.
Otherwise, the parameter is updated with a corrected stochastic gradient estimate of the form:
\begin{align*}
\widehat{\nabla}_{SVRG}f(\vtheta) = 
	\nabla f(\tilde{\vtheta}) + 
	\nabla g_{\vtheta}(x_i) -
	\nabla g_{\tilde{\vtheta}}(x_i),
\end{align*} 
where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$. The corrected gradient is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient. The resulting IFO complexity is $O(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
\section{Algorithm}
